---
title: '[''STAT 4610] HW-4'
author: "Michael Ghattas"
date: "9/24/2022"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Chapter - 4


## Problem - 13

```{r}
library(ISLR)
library(corrplot)
library(MASS)
library(class)
library(e1071)
```


### Part - (a)

```{r}
summary(Weekly)
corrplot(cor(Weekly[, -9]), method = "square")
```

-> Year and Volume are the variables that seem to have a significant linear relation. \


### Part - (b)

```{r}
Weekly.fit <- glm(Direction ~ Lag1 + Lag2 + Lag3 + Lag4 + Lag5 + Volume, data = Weekly, family = binomial)
summary(Weekly.fit)
```

-> Lag2 seems to be the only variable that has statistical significant at the level of significance. \


### Part - (c)

```{r}
logWeekly.prob = predict(Weekly.fit, type = 'response')
logWeekly.pred = rep("Down", length(logWeekly.prob))
logWeekly.pred[logWeekly.prob > 0.5] = "Up"

table(logWeekly.pred, Weekly$Direction)
```

-> The model predicted the weekly market trend correctly 56.11% of the time. \

-> $\frac{54 + 557}{54 + 48 + 430 + 557}$ = 0.5611 \

-> The model correctly predicted the Upward weekly trends 92.07% of the time. \

-> $\frac{557}{48 + 557}$ = 0.9207 \

-> The model correctly predicted the Downward weekly trends 11.15% of the time. \

-> $\frac{54}{54 + 430}$ = 0.1115 \


### part - (d)

```{r}
Direction = Weekly$Direction
train = (Weekly$Year < 2009)
test <- Weekly[!train, ]

Weekly.fit <- glm(Direction ~ Lag2, data = Weekly, family = binomial, subset = train)

logWeekly.prob = predict(Weekly.fit, test, type = "response")
logWeekly.pred = rep("Down", length(logWeekly.prob))
logWeekly.pred[logWeekly.prob > 0.5] = "Up"
Direction.test = Direction[!train]

table(logWeekly.pred, Direction.test)
mean(logWeekly.pred == Direction.test)
```

-> The model correctly predicted weekly trends at rate of 62.5% of the time. \

-> The model predicted upward trends 91.80% of the time. \

-> The model predicted downward trends 20.93%of the time. \


### part - (e)

```{r}
WeeklyLDA.fit <- lda(Direction ~ Lag2, data = Weekly, family = binomial, subset = train)
WeeklyLDA.pred <- predict(WeeklyLDA.fit, test)

table(WeeklyLDA.pred$class, Direction.test)
mean(WeeklyLDA.pred$class == Direction.test)
```

-> The Linear Discriminant Analysis (LDA) classifying model results are identical to the logistic regression model from part (e). \


### part - (f)

```{r}
WeeklyQDA.fit <- qda(Direction ~ Lag2, data = Weekly, subset = train)
WeeklyQDA.pred <- predict(WeeklyQDA.fit, test)

table(WeeklyQDA.pred$class, Direction.test)
mean(WeeklyQDA.pred$class == Direction.test)
```

-> The Quadratic Linear Analysis (QDA) model has 58.65% accuracy, which is lower than LDA, which has an accuracy of 62.5%. \

-> The QDA model only predicting the correctness of weekly upward trends while missing the downward weekly trends. \


### part - (g)

```{r}
Week.train = as.matrix(Weekly$Lag2[train])
Week.test = as.matrix(Weekly$Lag2[!train])
Direction.train = Direction[train]

set.seed(111)
WeekKNN.pred = knn(Week.train, Week.test, Direction.train, k = 1)

table(WeekKNN.pred, Direction.test)
mean(WeekKNN.pred == Direction.test)
```

-> The K-Nearest Neighbors (KNN) model resulted in a classifying model has ~51% accuracy. \

-> The KNN model has the lowest accuracy. \


### part - (h)

```{r}
WeeklyNB.fit <- naiveBayes(Direction ~ Lag2, data = Weekly, subset = train)
WeeklyNB.pred <- predict(WeeklyNB.fit, test)

table(WeeklyNB.pred, Direction.test)
mean(WeeklyNB.pred == Direction.test)
```

-> The Naive Bayes (NB) model has 58.65% accuracy, which is identical to the QDA model from part (f). \

-> Like the QDA model, the NB model also only predicting the correctness of weekly upward trends while missing the downward weekly trends. \


### part - (i)

-> Both the Logistic Regression model and LDA model have the best accuracy rate of 62.5%. \


### part - (j)

```{r}
#Logistic Regression with Lag2
Weekly.fit <- glm(Direction ~ Lag2, data = Weekly, family = binomial, subset = train)

logWeekly.prob = predict(Weekly.fit, test, type = "response")
logWeekly.pred = rep("Down", length(logWeekly.prob))
logWeekly.pred[logWeekly.prob > 0.5] = "Up"
Direction.test = Direction[!train]

table(logWeekly.pred, Direction.test)
mean(logWeekly.pred == Direction.test)


#LDA with Lag2
WeeklyLDA.fit <- lda(Direction ~ Lag2, data = Weekly,family = binomial, subset = train)
WeeklyLDA.pred <- predict(WeeklyLDA.fit, test)

table(WeeklyLDA.pred$class, Direction.test)
mean(WeeklyLDA.pred$class == Direction.test)


#QDA with with the 2nd power polynomial of Lag2
WeeklyQDA.fit = qda(Direction ~ poly(Lag2, 2), data = Weekly, subset = train)
WeeklyQDA.pred = predict(WeeklyQDA.fit, test)

table(WeeklyQDA.pred$class, Direction.test)
mean(WeeklyQDA.pred$class == Direction.test)


#KNN with Lag2 & K = 10
Week.train = as.matrix(Weekly$Lag2[train])
Week.test = as.matrix(Weekly$Lag2[!train])
Direction.train = Direction[train]

set.seed(222)
WeekKNN.pred = knn(Week.train, Week.test, Direction.train, k = 10)

table(WeekKNN.pred, Direction.test)
mean(WeekKNN.pred == Direction.test)


#KNN with Lag2 & K = 100
Week.train = as.matrix(Weekly$Lag2[train])
Week.test = as.matrix(Weekly$Lag2[!train])
Direction.train = Direction[train]

set.seed(222)
WeekKNN.pred = knn(Week.train, Week.test, Direction.train, k = 100)

table(WeekKNN.pred, Direction.test)
mean(WeekKNN.pred == Direction.test)
```

-> The Logistic Regression, LDA, and QDA(lag2^2) models have the best accuracy rate of 62.5%. \

-> While there where some improvement in accuracy with the KNN (k= 10, k = 100) models, their accuracy remains lower. \


# End.
