---
title: '[STAT 4610] HW-3 / Michael Ghattas'
author: "Michael Ghattas"
date: "9/18/2022"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Chapter - 4


## Problem - 13

```{r}
library(ISLR)
library(corrplot)
```

### Part - (a)

```{r}
summary(Weekly)
corrplot(cor(Weekly[,-9]), method="square")
```
-> Year and Volume are the variables that seem to have a significant linear relation. \


### Part - (b)

```{r}
Weekly.fit <- glm(Direction ~ Lag1 + Lag2 + Lag3 + Lag4 + Lag5 + Volume, data = Weekly, family = binomial)
summary(Weekly.fit)
```
-> Lag2 seems to be the only variable that has statistical significant at the level of significance. \


### Part - (c)

```{r}
logWeekly.prob = predict(Weekly.fit, type = 'response')
logWeekly.pred = rep("Down", length(logWeekly.prob))
logWeekly.pred[logWeekly.prob > 0.5] = "Up"

table(logWeekly.pred, Weekly$Direction)
```

-> The model predicted the weekly market trend correctly 56.11% of the time. \
-> $\frac{54 + 557}{54 + 48 + 430 + 557}$ = 0.5611 \
-> The model correctly predicted the Upward weekly trends 92.07% of the time. \
-> $\frac{557}{48 + 557}$ = 0.9207 \
-> The model correctly predicted the Downward weekly trends 11.15% of the time. \
-> $\frac{54}{54 + 430}$ = 0.1115 \


### part - (d)

```{r}
Direction = Weekly$Direction

train = (Weekly$Year < 2009)
test <- Weekly[!train, ]

Weekly.fit <- glm(Direction ~ Lag2, data = Weekly, family = binomial, subset = train)

logWeekly.prob = predict(Weekly.fit, test, type = "response")
logWeekly.pred = rep("Down", length(logWeekly.prob))
logWeekly.pred[logWeekly.prob > 0.5] = "Up"
Direction.test = Direction[!train]

table(logWeekly.pred, Direction.test)

mean(logWeekly.pred == Direction.test)
```

-> The model correctly predicted weekly trends at rate of 62.5% of the time. \
-> The model predicted upward trends 91.80% of the time. \
-> The model predicted downward trends 20.93%of the time. \


# End.