---
title: '[STAT 4610] HW-1 / Michael Ghattas'
author: "Michael Ghattas"
date: "9/6/2022"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Problem 10


## Start:

```{r}
library("ISLR")
head(Carseats)
```


### Part (a)

```{r}
lm.fit = lm(Sales ~ Price + Urban + US, data = Carseats)
summary(lm.fit)
```


### Part (b)

* If the price increases by USD 1000 while other predictors held constant, sales would decrease by ~54.46 units.
* A store location in relation to Urban areas has no affect on sales.
* US based stores will on average sell 1200 more carseats than international stores.


### Part (c)

* $Sales = \beta_{0} + \beta_{1} \cdot Price  + \beta_{2} \cdot UrbanYes + \beta_{3} \cdot USYes$


### Part (d)

* Urban: Its p-value is not statistically significant (p-value = 0.936).


### Part (e)

```{r}
lm.fit2 = lm(Sales ~ Price + US, data = Carseats)
summary(lm.fit2)
```


### Part (f)

* Given the R^2 values we can note that the two models are weak, with only ~23.5% of the data explained.

### Part (g)

```{r}
confint(lm.fit2)
```


### Part (h)

```{r}
par(mfrow = c(2, 2))
plot(lm.fit2)
```


* No treatment is needed, given the Normal Q-Q plot and the Residuals vs Leverage plot, we can note that outliers/high-leverage points are not significantly visible.


## End.